---
#title: "FEATURES_SELECTION1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 1. DATA PREPARATION

i)  Clean and Visualize features and relative stats
ii) Transform features: substitute values and normalization of absolute values
iii) Shuffle and split the data-set in training and test sets
iv) Save trainging and test data(cross validation will be performed directly on images)




```{r echo=FALSE}

## LOAD THE DATA

WD = '/Users/Remi/Dropbox/1_WORK/2015/4_PROJECTS/1_SOFTWARE/IMAGEJ_LIB/CELL SIGNALING ANALYSIS/CODE/2_CLASSIFY/BORDERCELLS/DATA_PREPARATION'

setwd(WD)

d0 <- read.delim("2_data_manual_selection.txt")

# exclude the text
d1 = d0[,3:length(d0)]



```


# 1.1 Clean data

```{r}
# total number of attributes/columns
DT = d1 # LOAD THE DATA IN 'd1'
num_cols = dim(DT)[2]

# report the columns with elements to correct
count_numeric_attributes=0
found_attributes = c('')
num_spec_values = c()

for(i in 1:num_cols){
    a = is.finite(DT[,i])
    l = length(a[a==FALSE])
    num_spec_values[i] = l
    if(l>0){
         count_numeric_attributes = count_numeric_attributes +1
        found_attributes[count_numeric_attributes] = names(DT[i])
        #print(names(DT[i]))
        #print(DT[,i][a==FALSE])
    }
}

if(count_numeric_attributes>0)print(found_attributes)

print(paste('Number of found attributes with special characters:', count_numeric_attributes))

```



# 1.2 Show features and statistics (log scale)
```{r}
d_edge = d1[,2:length(d1)]
d_edge_mtx =log(data.matrix(d_edge, rownames.force = NA),2)
op <- par(mar = c(0,10,0,0))
boxplot(d_edge_mtx, horizontal=TRUE, las=2, cex.axis = 0.7, ylim=c(0,20))
par(op)
```


# 1.3 Transform features:substitute nucleus and voronoi x,y position with their distance
```{r}
# Claculate the distance
Nx = d1$NUCL.dna.X
Ny = d1$NUCL.dna.Y
Vx = d1$VORO.protein3.CENTRE_X
Vy = d1$VORO.protein3.CENTRE_Y

dist= sqrt((Nx-Vx)^2+(Ny-Vy)^2)

# add distance column to the data frame
dd = d1
n_col = dim(dd)[2]
dd[,n_col+1] = dist
names(dd)[n_col+1]='NUCL_VORO.Distance'

# DROP from nuclei and VORO features: X, Y columns 
drops <- c("NUCL.dna.X","NUCL.dna.Y", "VORO.protein3.CENTRE_X", "VORO.protein3.CENTRE_Y")
d2 = dd[ , !(names(dd) %in% drops)]


# PLOT new features
dd = d2[,2:length(d2)]
dd =log(data.matrix(dd),2)
op <- par(mar = c(0,10,0,0))
boxplot(dd, horizontal=TRUE, las=2, cex.axis = 0.7, ylim=c(0,20))
par(op)


```



# 1.4 Transform features: normalize by nuclear dimension to make the classifier independent of absolute values

```{r}

# VORO.protein3.FERET
N_feret = d2$NUCL.dna.FERET
V_feret = d2$VORO.protein3.FERET
V_feret_NORM = V_feret/N_feret
#d1$VORO.protein3.AREA
N_area = d2$NUCL.dna.AREA
V_area = d2$VORO.protein3.AREA
V_area_NORM = V_area/N_area
#d1$CYTO.protein3.AREA
C_area = d2$CYTO.protein3.AREA
C_area_NORM = C_area/N_area
# NUCL_VORO.Distance
NV_dist = d2$NUCL_VORO.Distance
NV_dist_NORM = NV_dist/N_feret


# add VORO.Feret.norm
dd = d2
n_col = dim(dd)[2]
dd[,n_col+1] = V_feret_NORM
names(dd)[n_col+1]='VORO.Feret.norm'

# add VORO.Area.norm
n_col = dim(dd)[2]
dd[,n_col+1] = V_area_NORM
names(dd)[n_col+1]='VORO.Area.norm'

# add CYTO.Area.norm
n_col = dim(dd)[2]
dd[,n_col+1] = C_area_NORM
names(dd)[n_col+1]='CYTO.Area.norm'

# add NUCL_VORO.Distance
n_col = dim(dd)[2]
dd[,n_col+1] = NV_dist_NORM
names(dd)[n_col+1]='NUVO.distance.norm'


# DROP from nuclei and VORO features: X, Y columns 
drops <- c("NUCL.dna.FERET","VORO.protein3.AREA", "CYTO.protein3.AREA", "VORO.protein3.FERET", "NUCL.dna.AREA", "NUCL_VORO.Distance")
d3 = dd[ , !(names(dd) %in% drops)]


# SHOW the new features
dd = d3[,2:length(d3)]
dd =log(data.matrix(dd),2)
op <- par(mar = c(0,10,0,0))
boxplot(dd, horizontal=TRUE, las=2, cex.axis = 0.8, ylim=c(0,10))
par(op)

```



# 1.5 Shuffle and split the data-set in Training and Test sets
the cross validation will be performed directly on images (see Cross Validation folder)
```{r}

TRAIN_PERC = 0.75

## 75% of the sample size
smp_size <- floor(TRAIN_PERC * nrow(d3))

## set the seed to make your partition reproductible
set.seed(123)
train_ind <- sample(seq_len(nrow(d3)), size = smp_size)

train <- d3[train_ind, ]
test <- d3[-train_ind, ]

write.table(train, file = "4_data_training.txt", sep = "\t", quote=FALSE, row.names=FALSE, col.names = colnames(train))
write.table(test, file = "4_data_test.txt", sep = "\t", quote=FALSE, row.names=FALSE, col.names = colnames(test))


```
