
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 1. DATA PREPARATION 2: Normlization and Rescaling

i) Normalize distributions (e.g, log transformation)
ii) Rescaling features



```{r echo=FALSE}

## LOAD THE DATA

WD = '/Users/Remi/Dropbox/1_WORK/2015/4_PROJECTS/1_SOFTWARE/IMAGEJ_LIB/CELL SIGNALING ANALYSIS/CODE/2_CLASSIFY/BORDERCELLS/DATA_PREPARATION'

setwd(WD)

d0 <- read.delim("4_data_training.txt")

# exclude the text
d1 = d0[,1:length(d0)]



```


## 1.1 normal distribution: show histogram 

```{r}
par(mfrow=c(4,2))
par(mar = rep(2, 4))
dd_mtx =data.matrix(d1, rownames.force = NA)
n_col = dim(dd_mtx)[2]
for(i in 2:n_col){
  hist(dd_mtx[,i],100,main=names(d1)[i], cex.axis=1, font.main=1, cex.main=1.3)
}

```


## 1.2 normal distribution: Test log tansformation

```{r}
par(mfrow=c(4,2))
par(mar = rep(2, 4))
dd_mtx =log(data.matrix(d1, rownames.force = NA)+1,2)
n_col = dim(dd_mtx)[2]
for(i in 2:n_col){
  hist(dd_mtx[,i],100,main=names(d1)[i], cex.axis=1, font.main=1, cex.main=1.3)
}


```



## 1.3 normal distribution: Apply the log transformation
```{r}
d2 =d1
n_col = dim(dd_mtx)[2]
for(i in 2:n_col){
  a = d2[,i]
  d2[,i] =log(a+1,10)
}

par(mfrow=c(4,2))
par(mar = rep(2, 4))
dd_mtx =data.matrix(d2, rownames.force = NA)
n_col = dim(dd_mtx)[2]
for(i in 2:n_col){
  hist(dd_mtx[,i],50,main=names(d2)[i], cex.axis=1, font.main=1, cex.main=1.3)
}

```



## 1.4 rescaling (store the mean standard deviation for test and validation)
```{r}

dd = d2
FEATURES_MEAN = apply(data.matrix(dd),2,mean)
FEATURES_STDEV = apply(data.matrix(dd),2,sd)
for(i in 2:dim(dd)[2]){
  dd[,i]= (dd[,i] - FEATURES_MEAN[i])/ FEATURES_STDEV[i]
}

d3 = dd
dd =data.matrix(dd)
op <- par(mar = c(0,10,0,0))
boxplot(dd, horizontal=TRUE, las=2, cex.axis = 1, xlim=c(0,10))
par(op)
```

# 1.9 Save: transformed data set and mean and stdev features to be used for testing and cross validation 

```{r}

write.table(d3, file = "5_data_training_transf.txt", sep = "\t", quote=FALSE, row.names=FALSE, col.names = colnames(d3))
write.table(as.data.frame(FEATURES_MEAN) , file = "6_data_training_param_MEAN.txt", sep = "\t", quote=FALSE, row.names=T)
write.table(as.data.frame(FEATURES_STDEV), file = "7_data_training_param_STDEV.txt", sep = "\t", quote=FALSE, row.names=T)

```
